---
title: "Obtaining results"
output: html_document
date: "2025-05-23"
---

```{r}
library(tidyverse)
library(patchwork)
library(reshape2)
library(latex2exp)
theme_set(theme_bw())
```

# Dense covariance matrices

Import the results

```{r}
results <- readRDS("RDS/ALL_dense_XtX_2025.RDS")
table(results$algo)
results$algo <- ifelse(results$algo == "RIDGEWAY", "Ridgeway", results$algo)
```

- First, we make sure that all EP algorithm specifications provide very similar results

```{r}
eca <- results %>% filter(algo == "EP_Chol_algo")
ecb <- results %>% filter(algo == "EP_Chol_algoB")
eea <- results %>% filter(algo == "EP_Eigen_algoA")
eeb <- results %>% filter(algo == "EP_Eigen_algoB")
pairs(cbind(eca$log2prob, ecb$log2prob, eea$log2prob, eeb$log2prob))
pairs(cbind((ecb$log2prob-eca$log2prob)/eca$log2prob, 
            (eea$log2prob-eca$log2prob)/eca$log2prob, 
            (eeb$log2prob-eca$log2prob)/eca$log2prob ))
boxplot(cbind((ecb$log2prob-eca$log2prob), 
              (eea$log2prob-eca$log2prob), 
              (eeb$log2prob-eca$log2prob)))
TIM_dense <- cbind("CHOL A" = eca$time,
              "CHOL B" = ecb$time, 
              "EIG A" = eea$time, 
              "EIG B" = eeb$time)


DEP <- cbind("CA" = eca$log2prob, "CB" = ecb$log2prob, 
             "EA" = eea$log2prob, "EA" = eeb$log2prob)

REL_DELTA <- cbind("CHOL A vs CHOL B" = (eca$log2prob-ecb$log2prob)/ecb$log2prob, 
              "EIGEN A vs CHOL B" =(eea$log2prob-ecb$log2prob)/ecb$log2prob, 
              "EIGEN B vs CHOL B" =(eeb$log2prob-ecb$log2prob)/ecb$log2prob )

dense_EP <- REL_DELTA %>% melt() %>% 
ggplot()+
  geom_boxplot(aes(y = value,x=Var2))+
  xlab("")+ 
  theme(text = element_text(size=15))
dense_EP
```

- Given that the results are almost equivalent, we focus only on `EP_Chol_algoB`. We discard the other 3.

```{r}
results_epcB <- results %>% filter(algo != "EP_Eigen_algoA" , 
                               algo != "EP_Eigen_algoB" , 
                               algo != "EP_Chol_algo")
table(results_epcB$algo)
```

- To compute the errors, we use the `log2EP` prob as benchmark

```{r}
long_epcb    <- results_epcB %>% filter(algo == "EP_Chol_algoB") %>% pull(log2prob) %>% rep(4)
long_time    <- results_epcB %>% filter(algo == "EP_Chol_algoB") %>% pull(time) %>% rep(4)
long_bot_err <- results_epcB %>% filter(algo == "Botev") %>% pull(info) %>% as.numeric %>% rep(4)


d1 <- results_epcB %>% mutate(log2ep = long_epcb[1:4600], err = long_bot_err[1:4600]) %>%
  mutate(log2relative = (log2prob-log2ep)/log2ep, log2delta = (log2prob-log2ep),
         lowerr= (log2(2^log2prob*(1-err))-log2ep)/log2ep,
         upperr= (log2(2^log2prob*(1+err))-log2ep)/log2ep) %>% 
  group_by(b, dim, algo) %>% mutate(mlowerr=mean(lowerr, na.rm=TRUE),
                                 mupperr=mean(upperr, na.rm=TRUE)) %>% ungroup() %>% 
  filter(algo != "EP_Chol_algoB") %>% mutate(mlowerr2 = case_when(algo != "Botev"~NA,
                                                                  algo == "Botev"~mlowerr),
                                             mupperr2 = case_when(algo != "Botev"~NA,
                                                                  algo == "Botev"~mupperr)) %>% 
  mutate( dimf = paste("m =",dim))

d1$dimf <- factor(d1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                          "m = 512","m = 1024"))

plot1 = ggplot()+
  geom_hline(data=d1, aes(yintercept = 0),col=2)+
  geom_path(data=d1, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d1, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d1, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d1 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d1 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))


plot1
ggsave("NewFigures/DENSE_1024.pdf",width = 12, h=6)
ggsave("NewFigures/DENSE_1024_long.pdf",width = 15, h=6)


d11 <- d1 %>% filter( dim < 1024)
plot1 = ggplot()+
  geom_hline(data=d11, aes(yintercept = 0),col=2)+
  geom_path(data=d11, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d11, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d11, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d11 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d11 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
plot1
ggsave("NewFigures/DENSE_512.pdf",width = 10, h=6)
ggsave("NewFigures/DENSE_512_long.pdf",width = 15, h=6)


t1 <- results_epcB %>% mutate(timeep = long_time[1:4600]) %>%
  mutate(timerelative = (time)/timeep) %>%  
  filter(algo != "EP_Chol_algoB") %>% 
  mutate( dimf = paste("m =",dim))

t1$dimf <- factor(t1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                          "m = 512","m = 1024"))

time1 <- ggplot(t1 %>% filter(dim<1024))+
  geom_hline(aes(yintercept = 1),col=2)+
  geom_boxplot(aes(x = b,
                   y = timerelative, group = b),
               outlier.shape = NA)+
  facet_grid(algo~dimf)+
  ylab(TeX("Ratio between computational times"))+
  xlab(TeX("u"))+
  scale_y_log10()+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
time1
ggsave("NewFigures/DENSE_512_time.pdf",width = 10, h=6)
ggsave("NewFigures/DENSE_512_long_time.pdf",width = 15, h=6)
L <- ggplot(t1)+
  geom_hline(aes(yintercept = 1),col=2)+
  geom_boxplot(aes(x = b,
                   y = timerelative, group = b),
               outlier.shape = NA)+
  facet_grid(algo~dimf)+
  ylab(TeX("Ratio between computational times"))+
  xlab(TeX("u"))+
  scale_y_log10()+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
L
ggsave("NewFigures/DENSE_1024_time.pdf",width = 10, h=6)
ggsave("NewFigures/DENSE_1024_long_time.pdf",width = 15, h=6)

d11 <- d1 %>% filter( dim %in% c(16, 256, 512))
plot1 = ggplot()+
  geom_hline(data=d11, aes(yintercept = 0),col=2)+
  geom_path(data=d11, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d11, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d11, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d11 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d11 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
plot1
ggsave("NewFigures/DENSE_512_MAIN.pdf", width = 15, h=6)
```


## Probability curves

```{r}
d1 <- results_epcB %>% 
  mutate(log2prob2 = case_when(!is.finite(log2prob)~NA,
                               TRUE ~ log2prob)) %>%
  group_by(b, dim, algo) %>% mutate(ml2p = mean(log2prob2, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate( dimf = paste("m =",dim))

d1$dimf <- factor(d1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                                      "m = 512","m = 1024"))
d1 <- 
d1 %>% mutate(algo = case_when(algo == "EP_Chol_algoB"~ "EP Chol. Alg. 2",
                               TRUE ~ algo))
d11 <- d1
broken = d11 %>% filter(is.na(log2prob2)) %>% group_by(algo, dimf) %>% reframe(maxb = max(b))

ggplot()+
  geom_line(data=d11, aes(x=b,y=ml2p, col=algo))+
  geom_point(data = d11, aes(x = b,
                             y = log2prob2, group = b, col=algo), alpha=.2)+
  facet_wrap(~dimf, scale="free_y")+
  geom_point(data = d11 %>% filter(is.na(log2prob2)),
            aes(x=b,y=0, col=algo), pch="|")+
  geom_vline(data=broken,aes(xintercept = maxb,col=algo,lty=algo))+
  ylab(TeX("Estimated $log_2$ probabilities"))+
  xlab(TeX("u"))+
  scale_color_manual("Algorithm", values = c(1,2,3,4))+
  scale_linetype(guide=FALSE)+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1),
        legend.position = "bottom")
ggsave("NewFigures/DENSE_LINES_1024.pdf", width = 15, h=6)


ggplot()+
  geom_line(data=d11, aes(x=b,y=ml2p, col=algo))+
  geom_point(data = d11, aes(x = b,
                             y = log2prob2, group = b, col=algo), alpha=.2)+
  facet_wrap(~dimf, scale="free_y")+
  geom_point(data = d11 %>% filter(is.na(log2prob2)),
            aes(x=b,y=0, col=algo), pch="|")+
  geom_vline(data=broken,aes(xintercept = maxb,col=algo,lty=algo))+
  ylab(TeX("Estimated $log_2$ probabilities"))+
  xlab(TeX("u"))+
  scale_color_manual("Algorithm", values = c(1,2,3,4))+
  scale_linetype(guide=FALSE)+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1),
        legend.position = "none")
ggsave("NewFigures/DENSE_LINES_1024_noleg.pdf", width = 15, h=6)
```



# Fungible  covariance matrices

Import the results

```{r}
results <- readRDS("RDS/ALL_fungible_2025.RDS")
table(results$algo)
results$algo <- ifelse(results$algo == "RIDGEWAY", "Ridgeway", results$algo)
```

- First, we make sure that all EP algorithm specifications provide very similar results

```{r}
eca <- results %>% filter(algo == "EP_Chol_algo") # it is algo A
ecb <- results %>% filter(algo == "EP_Chol_algoB")
eea <- results %>% filter(algo == "EP_Eigen_algoA")
eeb <- results %>% filter(algo == "EP_Eigen_algoB")
pairs(cbind(eca$log2prob, ecb$log2prob, eea$log2prob, eeb$log2prob))
pairs(cbind((ecb$log2prob-eca$log2prob)/eca$log2prob, 
            (eea$log2prob-eca$log2prob)/eca$log2prob, 
            (eeb$log2prob-eca$log2prob)/eca$log2prob ))
boxplot(cbind((ecb$log2prob-eca$log2prob), 
              (eea$log2prob-eca$log2prob), 
              (eeb$log2prob-eca$log2prob)))
TIM_fung <- cbind("CHOL A" = eca$time,
              "CHOL B" = ecb$time, 
              "EIG A" = eea$time, 
              "EIG B" = eeb$time)


DEP <- cbind("CA" = eca$log2prob, "CB" = ecb$log2prob, 
             "EA" = eea$log2prob, "EA" = eeb$log2prob)

REL_DELTA <- cbind("CHOL A vs CHOL B" = (eca$log2prob-ecb$log2prob)/ecb$log2prob, 
                   "EIGEN A vs CHOL B" =(eea$log2prob-ecb$log2prob)/ecb$log2prob, 
               "EIGEN B vs CHOL B" =(eeb$log2prob-ecb$log2prob)/ecb$log2prob )

fung_EP <- REL_DELTA %>% melt() %>% 
ggplot()+
  geom_boxplot(aes(y = value,x=Var2))+
  xlab("")+ 
  theme(text = element_text(size=15))

fung_EP
```

- Given that the results are almost equivalent, we focus only on `EP_Chol_algoB`. We discard the other 3.

```{r}
results_epcB <- results %>% filter(algo != "EP_Eigen_algoA" , 
                               algo != "EP_Eigen_algoB" , 
                               algo != "EP_Chol_algo")
table(results_epcB$algo)
```

- To compute the errors, we use the `log2EP` prob as benchmark

```{r}
long_epcb    <- results_epcB %>% filter(algo == "EP_Chol_algoB") %>% pull(log2prob) %>% rep(4)
long_time    <- results_epcB %>% filter(algo == "EP_Chol_algoB") %>% pull(time) %>% rep(4)
long_bot_err <- results_epcB %>% filter(algo == "Botev") %>% pull(info) %>% as.numeric %>% rep(4)


d1 <- results_epcB %>% mutate(log2ep = long_epcb[1:4600], err = long_bot_err[1:4600]) %>%
  mutate(log2relative = (log2prob-log2ep)/log2ep, log2delta = (log2prob-log2ep),
         lowerr= (log2(2^log2prob*(1-err))-log2ep)/log2ep,
         upperr= (log2(2^log2prob*(1+err))-log2ep)/log2ep) %>% 
  group_by(b, dim, algo) %>% mutate(mlowerr=mean(lowerr, na.rm=TRUE),
                                 mupperr=mean(upperr, na.rm=TRUE)) %>% ungroup() %>% 
  filter(algo != "EP_Chol_algoB") %>% mutate(mlowerr2 = case_when(algo != "Botev"~NA,
                                                                  algo == "Botev"~mlowerr),
                                             mupperr2 = case_when(algo != "Botev"~NA,
                                                                  algo == "Botev"~mupperr)) %>% 
  mutate( dimf = paste("m =",dim))

d1$dimf <- factor(d1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                          "m = 512","m = 1024"))

plot1 = ggplot()+
  geom_hline(data=d1, aes(yintercept = 0),col=2)+
  geom_path(data=d1, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d1, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d1, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d1 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d1 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))


plot1
ggsave("NewFigures/FUNG_1024.pdf",width = 12, h=6)
ggsave("NewFigures/FUNG_1024_long.pdf",width = 15, h=6)


d11 <- d1 %>% filter( dim < 1024)
plot1 = ggplot()+
  geom_hline(data=d11, aes(yintercept = 0),col=2)+
  geom_path(data=d11, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d11, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d11, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d11 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d11 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
plot1
ggsave("NewFigures/FUNG_512.pdf",width = 10, h=6)
ggsave("NewFigures/FUNG_512_long.pdf",width = 15, h=6)


t1 <- results_epcB %>% mutate(timeep = long_time[1:4600]) %>%
  mutate(timerelative = (time)/timeep) %>%  
  filter(algo != "EP_Chol_algoB") %>% 
  mutate( dimf = paste("m =",dim))

t1$dimf <- factor(t1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                          "m = 512","m = 1024"))

time1 <- ggplot(t1 %>% filter(dim<1024))+
  geom_hline(aes(yintercept = 1),col=2)+
  geom_boxplot(aes(x = b,
                   y = timerelative, group = b),
               outlier.shape = NA)+
  facet_grid(algo~dimf)+
  ylab(TeX("Ratio between computational times"))+
  xlab(TeX("u"))+
  scale_y_log10()+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
time1
ggsave("NewFigures/FUNG_512_time.pdf",width = 10, h=6)
ggsave("NewFigures/FUNG_512_long_time.pdf",width = 15, h=6)

L <- ggplot(t1)+
  geom_hline(aes(yintercept = 1),col=2)+
  geom_boxplot(aes(x = b,
                   y = timerelative, group = b),
               outlier.shape = NA)+
  facet_grid(algo~dimf)+
  ylab(TeX("Ratio between computational times"))+
  xlab(TeX("u"))+
  scale_y_log10()+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
L
ggsave("NewFigures/FUNG_1024_time.pdf",width = 10, h=6)
ggsave("NewFigures/FUNG_1024_long_time.pdf",width = 15, h=6)


d11 <- d1 %>% filter( dim %in% c(16, 256, 512))
plot1 = ggplot()+
  geom_hline(data=d11, aes(yintercept = 0),col=2)+
  geom_path(data=d11, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d11, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d11, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d11 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d11 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
plot1
ggsave("NewFigures/FUNG_512_MAIN.pdf", width = 15, h=6)
```

## Probability curves

```{r}
d1 <- results_epcB %>% 
  mutate(log2prob2 = case_when(!is.finite(log2prob)~NA,
                               TRUE ~ log2prob)) %>%
  group_by(b, dim, algo) %>% mutate(ml2p = mean(log2prob2, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate( dimf = paste("m =",dim))

d1$dimf <- factor(d1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                                      "m = 512","m = 1024"))
d1 <- 
d1 %>% mutate(algo = case_when(algo == "EP_Chol_algoB"~ "EP Chol. Alg. 2",
                               TRUE ~ algo))
d11 <- d1
broken = d11 %>% filter(is.na(log2prob2)) %>% group_by(algo, dimf) %>% reframe(maxb = max(b))

ggplot()+
  geom_line(data=d11, aes(x=b,y=ml2p, col=algo))+
  geom_point(data = d11, aes(x = b,
                             y = log2prob2, group = b, col=algo), alpha=.2)+
  facet_wrap(~dimf, scale="free_y")+
  geom_point(data = d11 %>% filter(is.na(log2prob2)),
            aes(x=b,y=0, col=algo), pch="|")+
  geom_vline(data=broken,aes(xintercept = maxb,col=algo,lty=algo))+
  ylab(TeX("Estimated $log_2$ probabilities"))+
  xlab(TeX("u"))+
  scale_color_manual("Algorithm", values = c(1,2,3,4))+
  scale_linetype(guide=FALSE)+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1),
        legend.position = "bottom")
ggsave("NewFigures/FUNG_LINES_1024.pdf", width = 15, h=6)

ggplot()+
  geom_line(data=d11, aes(x=b,y=ml2p, col=algo))+
  geom_point(data = d11, aes(x = b,
                             y = log2prob2, group = b, col=algo), alpha=.2)+
  facet_wrap(~dimf, scale="free_y")+
  geom_point(data = d11 %>% filter(is.na(log2prob2)),
            aes(x=b,y=0, col=algo), pch="|")+
  geom_vline(data=broken,aes(xintercept = maxb,col=algo,lty=algo))+
  ylab(TeX("Estimated $log_2$ probabilities"))+
  xlab(TeX("u"))+
  scale_color_manual("Algorithm", values = c(1,2,3,4))+
  scale_linetype(guide=FALSE)+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1),
        legend.position = "none")
ggsave("NewFigures/FUNG_LINES_1024_noleg.pdf", width = 15, h=6)
```



# Constant correlation matrix

```{r}
results <- readRDS("RDS/ALL_rhofixed_2025.RDS")
truth   <- readRDS("RDS/00_trueLOGProbs_constantSigma.RDS")
table(results$algo)
results$algo <- ifelse(results$algo == "RIDGEWAY", "Ridgeway", results$algo)
```

- First, we make sure that all EP algorithm specifications provide very similar results

```{r}
eca <- results %>% filter(algo == "EP_Chol_algoA")
ecb <- results %>% filter(algo == "EP_Chol_algoB")
eea <- results %>% filter(algo == "EP_Eigen_algoA")
eeb <- results %>% filter(algo == "EP_Eigen_algoB")
pairs(cbind(eca$log2prob, ecb$log2prob, eea$log2prob, eeb$log2prob))
pairs(cbind((ecb$log2prob-eca$log2prob)/eca$log2prob, 
            (eea$log2prob-eca$log2prob)/eca$log2prob, 
            (eeb$log2prob-eca$log2prob)/eca$log2prob ))

boxplot(cbind((ecb$log2prob-eca$log2prob), 
              (eea$log2prob-eca$log2prob), 
              (eeb$log2prob-eca$log2prob)))
TIM_const <- cbind("CHOL A" = eca$time,
              "CHOL B" = ecb$time, 
              "EIG A" = eea$time, 
              "EIG B" = eeb$time)

DEP <- cbind("CA" = eca$log2prob, "CB" = ecb$log2prob, 
             "EA" = eea$log2prob, "EA" = eeb$log2prob)

REL_DELTA <- cbind("CHOL A vs CHOL B" = (eca$log2prob-ecb$log2prob)/ecb$log2prob, 
              "EIGEN A vs CHOL B" =(eea$log2prob-ecb$log2prob)/ecb$log2prob, 
              "EIGEN B vs CHOL B" =(eeb$log2prob-ecb$log2prob)/ecb$log2prob )

const_EP <- 
REL_DELTA %>% melt() %>% 
ggplot()+
  geom_boxplot(aes(y = value,x=Var2))+
  xlab("")+ 
  theme(text = element_text(size=15))

const_EP

pp1 <- (dense_EP+coord_flip()+fung_EP+coord_flip()+const_EP+coord_flip())
pp1

q1 = dense_EP$data %>% as_tibble() %>% mutate(type = "Dense Cov.")
q2 = fung_EP$data %>% as_tibble() %>% mutate(type = "Fungible Cov.")
q3 = const_EP$data %>% as_tibble() %>% mutate(type = "Constant Cov.")
all_EP <- bind_rows(q1,q2,q3)
table(all_EP$Var2)
all_EP <- all_EP %>% mutate(Var3 = case_when(Var2 == "CHOL A vs CHOL B" ~ "Chol. Alg. 1 vs\nChol. Alg. 2",
                                             Var2 == "EIGEN A vs CHOL B" ~ "Eig. Alg. 1 vs\nChol. Alg. 2",
                                             Var2 == "EIGEN B vs CHOL B" ~ "Eig. Alg. 2 vs\nChol. Alg. 2"))

plotEP <- all_EP %>% as_tibble() %>% 
  ggplot()+facet_wrap(~type)+
  geom_hline(yintercept = 0,col=2)+
  geom_boxplot(aes(x = Var3, y = value))+
  xlab("") + ylab(TeX("Relative difference in $log_2$ probabilities"))+
    theme(text = element_text(size=15))#+coord_flip()
plotEP
ggsave(filename = "NewFigures/logratio_prob_EP.pdf",h=5,w=15)
```

Same for time

```{r}
q1 = TIM_dense %>% as_tibble() %>% mutate(type = "Dense Cov.")
q2 = TIM_fung %>% as_tibble() %>% mutate(type = "Fungible Cov.")
q3 = TIM_const %>% as_tibble() %>% mutate(type = "Constant Cov.")
alltime_EP <- bind_rows(q1,q2,q3)
Meltime <- melt(alltime_EP,"type")
table(Meltime$variable)
all_EP <- Meltime %>% mutate(Var3 = case_when(variable == "CHOL A" ~ "Chol. Alg. 1",
                                              variable == "CHOL B" ~ "Chol. Alg. 2",
                                              variable ==  "EIG A" ~ "Eig. Alg. 1",
                                              variable ==  "EIG B" ~ "Eig. Alg. 2"))

plotEP <- all_EP %>% as_tibble() %>% 
  ggplot()+facet_wrap(~type)+
  geom_hline(yintercept = 0,col=2)+
  geom_boxplot(aes(x = Var3, y = value))+
  xlab("") + ylab(TeX("Seconds"))+
    theme(text = element_text(size=15))#+coord_flip()
plotEP
ggsave("NewFigures/times_EP.pdf",h=5,w=15)
```

```{r}
q1 = TIM_dense %>% as_tibble() %>% mutate(type = "Dense Cov.") %>% mutate(
  CACB = `CHOL A`/`CHOL B`,
  EACB = `EIG A`/`CHOL B`,EBCB = `EIG B`/`CHOL B`
)
q2 = TIM_fung %>% as_tibble() %>% mutate(type = "Fungible Cov.") %>% mutate(
  CACB = `CHOL A`/`CHOL B`,
  EACB = `EIG A`/`CHOL B`,EBCB = `EIG B`/`CHOL B`
)
q3 = TIM_const %>% as_tibble() %>% mutate(type = "Constant Cov.") %>% mutate(
  CACB = `CHOL A`/`CHOL B`,
  EACB = `EIG A`/`CHOL B`,EBCB = `EIG B`/`CHOL B`
)

alltime_EP <- bind_rows(q1,q2,q3) %>% select(type,CACB,EACB,EBCB)
Meltime <- melt(alltime_EP,"type")
table(Meltime$variable)
all_EP <- Meltime %>% mutate(Var3 = case_when(variable == "CACB" ~ "Chol. Alg. 1/\nChol. Alg. 2",
                                              variable ==  "EACB" ~ "Eig. Alg. 1/\nChol. Alg. 2",
                                              variable ==  "EBCB" ~ "Eig. Alg. 2/\nChol. Alg. 2"))
plotEP <- all_EP %>% as_tibble() %>% 
  ggplot()+facet_wrap(~type)+
  geom_hline(yintercept = 1,col=2)+
  geom_boxplot(aes(x = Var3, y = value))+
  xlab("") + scale_y_log10()+
    ylab(TeX("Ratio between computational times"))+
    theme(text = element_text(size=15))#+coord_flip()
plotEP
ggsave("NewFigures/ratio_of_times_EP_log10.pdf",h=5,w=15)

plotEP <- all_EP %>% as_tibble() %>% 
  ggplot()+facet_wrap(~type)+
  geom_hline(yintercept = 1,col=2)+
  geom_boxplot(aes(x = Var3, y = value))+
  xlab("") + 
    ylab(TeX("Ratio between computational times"))+
    theme(text = element_text(size=15))#+coord_flip()
plotEP
ggsave("NewFigures/ratio_of_times_EP.pdf",h=5,w=15)
```


- Given that the results are almost equivalent, we focus only on `EP_Chol_algoB`. We discard the other 3.

```{r}
results_epcB <- results %>% filter(algo != "EP_Eigen_algoA" , 
                                   algo != "EP_Eigen_algoB" , 
                                   algo != "EP_Chol_algoA")
table(results_epcB$algo)
results_epcB <- results_epcB %>% left_join(truth %>% rename(dim = Ns),by = c("b","dim","rho")) %>% mutate(log2true = (Truep) / log(2))
table(results_epcB$algo)
sum(is.na(results_epcB$log2true))
```

- To compute the errors, we use the `log2EP` prob as benchmark

```{r}
long_time    <- results_epcB %>% filter(algo == "EP_Chol_algoB") %>% pull(time) %>% rep(4)

d1 <- results_epcB %>%
  mutate(log2relative = (log2prob-log2true)/log2true, 
         log2delta = (log2prob-log2true),
         err = as.numeric(info),
         lowerr= (log2(2^log2prob*(1-err))-log2true)/log2true,
         upperr= (log2(2^log2prob*(1+err))-log2true)/log2true) %>% 
  group_by(algo, rho, dim, b) %>% 
  mutate(mlowerr=mean(lowerr, na.rm=TRUE),
         mupperr=mean(upperr, na.rm=TRUE)) %>% 
  ungroup() %>% mutate(mlowerr2 = case_when(algo != "Botev"~NA,
                                                                  algo == "Botev"~mlowerr),
                                             mupperr2 = case_when(algo != "Botev"~NA,
                                                                  algo == "Botev"~mupperr)) %>% 
  mutate( dimf = paste("m =",dim))

d1 <- d1 %>% mutate(algo = case_when(algo == "EP_Chol_algoB"~"EP Chol. Alg. 2", TRUE ~ algo))
d1$dimf <- factor(d1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                          "m = 512","m = 1024"))
d1$algo <- factor(d1$algo,levels =  c("Botev","EP Chol. Alg. 2","TLRank","Ridgeway"))


d0 = d1

plots = function(d0,rho_val){
  L = list()
  d11 <- d0 %>% filter(rho==rho_val) 
L[[1]] <- 
  
  ggplot()+
  geom_hline(data=d11, aes(yintercept = 0),col=2)+
  geom_path(data=d11, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d11, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d11, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d11 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d11 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))

d11 = d11 %>% filter(dim<1024)
L[[2]] <-   
  ggplot()+
  geom_hline(data=d11, aes(yintercept = 0),col=2)+
  geom_path(data=d11, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d11, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d11, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d11 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d11 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))

d11 = d11 %>% filter(dim %in% c(16, 256, 512))
L[[3]] <-   
  ggplot()+
  geom_hline(data=d11, aes(yintercept = 0),col=2)+
  geom_path(data=d11, aes(x=b, y = c(mlowerr2)),col=4)+
  geom_path(data=d11, aes(x=b, y = c(mupperr2)),col=4)+
  geom_boxplot(data=d11, aes(x = b,
                   y = log2relative, group = b),outlier.shape = NA)+
  geom_point(data=d11 %>% filter(is.na(log2prob)),aes(x=b,y=0),pch="x",col="darkred")+
  geom_point(data=d11 %>% filter((log2prob)==-Inf),aes(x=b,y=0),pch="|",col="darkred")+
  facet_grid(algo~dimf)+
  ylab(TeX("Relative difference in $log_2$ probabilities"))+
  xlab(TeX("u"))+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
return(L)
}


l00 = plots(d1,00)
l00[[1]]
ggsave("NewFigures/Const00_1024.pdf",width = 12, h=6)
ggsave("NewFigures/Const00_1024_long.pdf",width = 15, h=6)
l00[[2]]
ggsave("NewFigures/Const00_512.pdf",width = 10, h=6)
ggsave("NewFigures/Const00_512_long.pdf",width = 15, h=6)
l00[[3]]
ggsave("NewFigures/Const00_512_MAIN.pdf",h = 9, w=15)

l25 = plots(d1,.25)
l25[[1]]
ggsave("NewFigures/Const25_1024.pdf",width = 10, h=6)
ggsave("NewFigures/Const25_long_1024.pdf",width = 15, h=6)
l25[[2]]
ggsave("NewFigures/Const25_512.pdf",width = 10, h=6)
ggsave("NewFigures/Const25_512_long.pdf",width = 15, h=6)
l25[[3]]
ggsave("NewFigures/Const25_512_MAIN.pdf",h = 9, w=15)

l50 = plots(d1,.50)
l50[[1]]
ggsave("NewFigures/Const50_1024.pdf",width = 10, h=6)
ggsave("NewFigures/Const50_1024_long.pdf",width = 15, h=6)
l50[[2]]
ggsave("NewFigures/Const50_512.pdf",width = 10, h=6)
ggsave("NewFigures/Const50_512_long.pdf",width = 15, h=6)
l50[[3]]
ggsave("NewFigures/Const50_512_MAIN.pdf",h = 9, w=15)



l75 = plots(d1,.75)
l75[[1]]
ggsave("NewFigures/Const75_1024.pdf",h = 9, w=15)
ggsave("NewFigures/Const75_1024_long.pdf",width = 15, h=6)
l75[[2]]
ggsave("NewFigures/Const75_512.pdf",width = 10, h=6)
ggsave("NewFigures/Const75_512_long.pdf",width = 15, h=6)
l75[[3]]
ggsave("NewFigures/Const75_512_MAIN.pdf",width = 9, h=15)
```


```{r}
t1 <- results_epcB %>% mutate(timeep = long_time[1:18400]) %>%
  mutate(timerelative = (time)/timeep) %>%  
  filter(algo != "EP_Chol_algoB") %>% 
  mutate( dimf = paste("m =",dim))

table(t1$algo)
t1$algo <- factor(t1$algo,levels =  c("Botev", "TLRank", "Ridgeway"))


t1$dimf <- factor(t1$dimf,levels =  c("m = 16","m = 64","m = 128","m = 256",
                          "m = 512","m = 1024"))

L <- ggplot(t1 %>% filter(dim<1024))+
  geom_hline(aes(yintercept = 1),col=2)+
  geom_boxplot(aes(x = b,
                   y = timerelative, group = b),
               outlier.shape = NA)+
  facet_grid(algo~dimf)+
  ylab(TeX("Ratio between computational times"))+
  xlab(TeX("u"))+
  scale_y_log10()+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
L
ggsave("NewFigures/ConstALL_512_time.pdf",width = 10, h=6)
ggsave("NewFigures/ConstALL_512_long_time.pdf",width = 15, h=6)

L <- ggplot(t1)+
  geom_hline(aes(yintercept = 1),col=2)+
  geom_boxplot(aes(x = b,
                   y = timerelative, group = b),
               outlier.shape = NA)+
  facet_grid(algo~dimf)+
  ylab(TeX("Ratio between computational times"))+
  xlab(TeX("u"))+
  scale_y_log10()+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1))
L
ggsave("NewFigures/ConstALL_1024_time.pdf",width = 10, h=6)
ggsave("NewFigures/ConstALL_1024_long_time.pdf",width = 15, h=6)
```


## Probability curves

```{r}
d2 <- results_epcB %>% 
  mutate(log2prob2 = case_when(!is.finite(log2prob)~NA,
                               TRUE ~ log2prob)) %>%
  group_by(b, rho,#######
           dim, algo) %>% mutate(ml2p = mean(log2prob2, na.rm=TRUE)) %>% 
  ungroup() %>% 
  mutate( dimf = paste("m ==",dim),
          rhof = paste0("rho ==",rho),
          )
d2 <- 
d2 %>% mutate(algo = case_when(algo == "EP_Chol_algoB"~ "EP Chol. Alg. 2",
                               TRUE ~ algo))

d2$dimf <- factor(d2$dimf,levels =  c("m == 16","m == 64","m == 128","m == 256",
                                      "m == 512","m == 1024"))
d2$rhof <- factor(d2$rhof)#,levels =  c("$\\rho = 0","m = 64","m = 128","m = 256",
                           #           "m = 512","m = 1024"))

table(d2$algo)
d2$algo <- factor(d2$algo, levels = c("Botev", 
                                      "EP Chol. Alg. 2",   
                                      "TLRank" ,     "Ridgeway"))

broken <- d2 %>% filter(is.na(log2prob2)) %>% group_by(algo, rhof, dimf) %>% 
  reframe(maxb = max(b)) 

ggplot()+
  geom_line(data=d2, aes(x=b,y=ml2p, col=algo)) +
  geom_point(data = d2, aes(x = b,
                             y = log2prob2, group = b, col=algo), alpha=.2) +
  facet_grid(rhof~dimf, scale="free_y", label = label_parsed) +
  geom_point(data = d2 %>% filter(is.na(log2prob2)),
            aes(x=b,y=0, col=algo), pch="|") +
  geom_vline(data=broken,aes(xintercept = maxb,col=algo,lty=algo))+
  ylab(TeX("Estimated $log_2$ probabilities"))+
  xlab(TeX("u"))+
  scale_color_manual("Algorithm", values = c(1,2,3,4))+
  scale_linetype(guide=FALSE)+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1),
        legend.position = "bottom")
ggsave("NewFigures/RHO_LINES_1024.pdf", width = 15, h=9)



ggplot(d2)+
  geom_line(data=d2, aes(x=b,y=ml2p, col=algo))+
  facet_grid( rhof ~ dimf, scale="free_y", label = label_parsed)+
  geom_point(data = d2, aes(x = b,
                             y = log2prob2, group = b, col=algo), alpha=.2)+
  geom_point(data = d2 %>% filter(is.na(log2prob2)),
            aes(x=b,y=0, col=algo), pch="|")+
  geom_vline(data=broken,aes(xintercept = maxb,
                             col=algo, lty = algo))+
  ylab(TeX("Estimated $log_2$ probabilities"))+
  xlab(TeX("u"))+
  scale_color_manual("Algorithm", values = c(1,2,3,4))+
  scale_linetype(guide=FALSE)+
  theme(text = element_text(size=15), 
        axis.text.x = element_text(angle = 00, hjust = 1),
        legend.position = "none")
ggsave("NewFigures/RHO_LINES_1024_noleg.pdf", width = 15, h=9)
```

