---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# EPmvnCDF Repository

<!-- 
[![Last Commit](https://img.shields.io/github/last-commit/fradenti/EPmvnCDF)](https://github.com/fradenti/EPmvnCDF)
 -->

The repository **EPmvnCDF**, short for  

*Expectation propagation for the estimation of multivariate normal cumulative distribution functions*
<!--**DUA**l probit marginal **L**ikelihood **E**stimation via **EP** **A**pproximation, -->

contains the R package implementing the estimation strategy proposed in 

*''Multivariate Gaussian cumulative distribution functions as the marginal likelihood of their dual Bayesian probit model''*   
by A. Fasano and F. Denti

and the code needed to reproduce the simulation studies and the figures.

First, set the working directory to this directory - you can open the R project `EPmvnCDF.Rproj`. Then, install the R package `EPmvnCDF`. You can either use the built versions via

```{r, eval = FALSE}
install.packages("EPmvnCDF_0.1.0.tar.gz", type = "source")
```

or open the `R` project contained in the `EPmvnCDF/` folder and build it from there.

Then you can run the code.

In particular, 

- Scripts starting with letter `A` contain useful functions;
- Scripts starting with letter `B` contain the code to run the four methods. Results will be stored in the corresponding subfolder into the `RDS/` folder;
- The script starting with letter `C` contains the code to extract the results from the saved `.RDS` files. Data frames will be created created inside `RDS/`; 
- Scripts starting with letter `D` contain the code to produce the plots showed in the paper. They will be saved in the folder `OUTPUT/`.

